---
title: "scouts"
format: html
---

## About

This page summarizes some data cleaning and analyses we did in the Gilmore lab meeting on 2026-04-08.

## Setup

The raw data file is located in `include/csv'. The file name is `TrainedLeader_JVC_497_20260408.csv`.

::: {.callout-important}
This file contains names and other identifiable information. So, I have added it to a `.gitignore` file so that it is not tracked.
:::

We load some basic functions.

```{r}
library(tidyverse)
library(ggplot2)
library(dplyr)
```

## Import

We visually examined the data file first.
This file had header rows. We deleted them and saved a copy as `scouts-no-header.csv`.

Now, we import the file.

```{r}
#| label: import-scouts

scouts_df <- readr::read_csv(file = "include/csv/scouts-no-header.csv")
```

```{r}
str(scouts_df)
```

## Questions

1. Who is not trained?

```{r}
not_trained_df <- scouts_df |>
  dplyr::filter(Trained == "NO") |>
  dplyr::arrange(MemberID) |>
  dplyr::select(-c("Council", "Service_Area", "Sub_District", "Gender_Accepted", "Chartered_Org_Name", "Zip_Code"))
```

```{r}
not_trained_df |>
  kableExtra::kbl()
```

How many positions per person?

```{r}
not_trained_count_df <- not_trained_df |>
  dplyr::group_by(MemberID, Last_Name, First_Name, Email) |>
  dplyr::summarise(n_pos = dplyr::n()) |>
  dplyr::arrange(desc(n_pos)) 

readr::write_csv(not_trained_count_df, file = "include/csv/not-trained-count.csv")

not_trained_count_df |>
  kableExtra::kbl()

not_trained_count_df |>
  ggplot() +
  aes(x = n_pos) +
  geom_histogram() +
  ggtitle("Histogram of leaders who are missing trainings.")
```

