---
title: "Discussion of Hadad et al. 2015"
format: html
---

## About

This document provides supporting information for the lab group discussion of @Hadad2015-qd.

## Setup

In order to generate images that support the concepts here, we set up a Python environment.

First, we install Python. We only run this once.

```{r}
#| eval: false
install.packages("reticulate")
library(reticulate)

# Installs a standalone Python version managed by reticulate (works cross-platform)
reticulate::install_python(version = "3.12")
reticulate::virtualenv_create(
  envname = "my_python_env",
  python = "3.12" 
)
reticulate::virtualenv_install(
  envname = "my_python_env", 
  packages = c("numpy", "matplotlib", "pandas", "scikit-learn")
)
```

Each time we render the file, we use the virtual environment we created in the previous step.

```{r}
reticulate::use_virtualenv("my_python_env", required = TRUE)
```

## Core question(s)

- Motion perception relates to clinical and development disorders.
- Motion perception/processing affects other aspect of perception.
- How does motion perception develop?
- When is it adult-like?
- What brain systems process motion?

## Key concepts

- Local vs. global motion
- Grating stimuli

```{python}
# Generated from Google AI query 2026-03-04
import numpy as np
import matplotlib.pyplot as plt
import math

def generate_sinusoidal_grating(size, spatial_frequency, angle=0, phase=0):
    """
    Generates a sinusoidal grating.

    Args:
        size (int): The width and height of the square grating in pixels.
        spatial_frequency (float): The spatial frequency (cycles per pixel, 
                                   or inverse of wavelength in pixels).
        angle (float): The orientation of the grating in degrees.
        phase (float): The phase of the grating in degrees.

    Returns:
        numpy.ndarray: A 2D array representing the grating.
    """
    # Create a grid of coordinates
    x = np.arange(size)
    y = np.arange(size)
    X, Y = np.meshgrid(x, y)

    # Convert angle and phase to radians
    angle_rad = math.radians(angle)
    phase_rad = math.radians(phase)

    # Calculate the gradient along the specified angle
    # This creates a ramp that changes value along the desired orientation
    gradient = X * math.cos(angle_rad) + Y * math.sin(angle_rad)

    # Apply the sine wave function to the gradient
    # The result will range between -1 and 1
    grating = np.sin((2 * math.pi * gradient * spatial_frequency) + phase_rad)

    return grating

# --- Example Usage ---
# Parameters
grating_size = 512 # pixels
frequency = 0.02 # cycles per pixel (e.g. 1 cycle every 50 pixels)
orientation = 45 # degrees
start_phase = 90 # degrees

# Generate the grating
grating_pattern = generate_sinusoidal_grating(
    grating_size,
    frequency,
    orientation,
    start_phase
)

# Display the grating using Matplotlib
plt.imshow(grating_pattern, cmap='gray', vmin=-1, vmax=1)
plt.title(f"Sinusoidal Grating at {orientation}°")
plt.axis('off') # Hide axes
plt.show()
```

  - A moving grating can yield a Barber Pole illusion:
  
  ![@Wikipedia-contributors2025-hv^["By The original uploader was Rokers at English Wikipedia. - Transferred from en.wikipedia to Commons., CC BY-SA 3.0, https://commons.wikimedia.org/w/index.php?curid=1881350"]](https://upload.wikimedia.org/wikipedia/commons/3/39/Barberpole_illusion_animated.gif)
  
  - Plaids
  
::: {#fig-plaids-ambiguous-not}
<iframe width="560" height="315" src="https://www.youtube.com/embed/WhbZesV2GRk?si=TeA4BPTaGS7Dz6bs" title="YouTube video player" frameborder="0" allow="accelerometer; autoplay; clipboard-write; encrypted-media; gyroscope; picture-in-picture; web-share" referrerpolicy="strict-origin-when-cross-origin" allowfullscreen></iframe>

@ricardomartins2016-pt
:::

- Dot stimuli
  - Random Dot Kinematograms (RDKs)
    - [Example](https://databrary.org/volume/49) from Databrary

- Gabor stimuli


  
